\\pkg{casebase}: An Alternative Framework For Survival Analysis and Comparison of Event Rates

### 2021-10-13 analysis/paper/bhatnagar-turgeon-islam-hanley-saarela.Rmd
load ggplot2 in setup.

options(kableExtra.latex.load_packages = FALSE)
library(casebase)
library(colorspace)
library(cowplot)
library(data.table)
library(glmnet)
library(kableExtra)
library(magrittr)
library(pracma)
library(flexsurv) 
library(prodlim)
library(riskRegression)
library(splines)
library(survival)
library(tibble)
library(visreg)
library(lubridate)
library(tidyverse)
library(scales)
options(digits = 2)
# set the seed for reproducible output
set.seed(1234)

# Turn on/off sections for faster compilation
eval_introduction <- TRUE
eval_theory <- TRUE
eval_implementation <- TRUE
eval_cs1 <- TRUE
eval_cs2 <- TRUE
eval_cs3 <- TRUE
eval_cs4 <- FALSE
eval_colophon <- FALSE # Only for draft
echo_plot_code <- TRUE
# Define theme for plots
paper_gg_theme <- theme_bw(base_size = 10) + theme(legend.position = "bottom")
# Define colour palette for CS 3
q4 <- qualitative_hcl(4, palette = "Dark 3")
names(q4) <- c("Cox", "Pen. Cox", "Pen. CB", "K-M")
```

@@

Several authors have since pursued fully parametric approaches that made the 
fitting of smooth survival curves much more transparent and intuitive through 
generalized linear models. The key feature of these procedures is splitting the 
time axis into discrete intervals. Whitehead [-@whitehead1980fitting] showed the 
equivalence between a Cox model and a Poisson regression with a parameter for 
each event time; Carstensen [-@copenhagen2012needs] provides a nice exposition of this 
equivalence with a real data example and supporting R code for computing 
standard errors. Arjas & Haara [-@arjas1987logistic] and Efron [-@efron1988logistic] treated each 
patient-day as a Bernoulli random variable with probability equal to the 
discrete hazard rate. A potential issue with these approaches is that the 
number of time bins need to be chosen by the data analyst. On the one hand, a fine grouping 
of times may result in few (or none) events per interval, which then leads 
to instability in the Newton-Raphson procedure for estimation 
[@kalbfleisch2011statistical]. On the other hand, a coarse grouping could potentially mask 
nonlinear trends in the hazard function.

>>>>

options(kableExtra.latex.load_packages = FALSE)
library(casebase)
library(ggplot2)
library(colorspace)
library(cowplot)
library(data.table)
library(glmnet)
library(kableExtra)
library(magrittr)
library(pracma)
library(flexsurv) 
library(prodlim)
library(riskRegression)
library(splines)
library(survival)
library(tibble)
library(visreg)
library(lubridate)
library(tidyverse)
library(scales)
options(digits = 2)
# set the seed for reproducible output
set.seed(1234)

# Turn on/off sections for faster compilation
eval_introduction <- TRUE
eval_theory <- TRUE
eval_implementation <- TRUE
eval_cs1 <- TRUE
eval_cs2 <- TRUE
eval_cs3 <- TRUE
eval_cs4 <- FALSE
eval_colophon <- FALSE # Only for draft
echo_plot_code <- TRUE
# Define theme for plots
paper_gg_theme <- ggplot2::theme_bw(base_size = 10) + ggplot2::theme(legend.position = "bottom")
# Define colour palette for CS 3
q4 <- colorspace::qualitative_hcl(4, palette = "Dark 3")
names(q4) <- c("Cox", "Pen. Cox", "Pen. CB", "K-M")
```

@@

Several authors have since pursued fully parametric approaches that made the 
fitting of smooth survival curves much more transparent and intuitive through 
generalized linear models. The key feature of these procedures is splitting the 
time axis into discrete intervals. Whitehead [-@whitehead1980fitting] showed the 
equivalence between a Cox model and a Poisson regression with a parameter for 
each event time; Carstensen [-@copenhagen2012needs] provides a nice exposition of this 
equivalence with a real data example and supporting R code for computing 
standard errors. Arjas & Haara [-@arjas1987logistic] and Efron [-@efron1988logistic] treated each 
patient-day as a Bernoulli random variable with probability equal to the 
discrete hazard rate. A potential issue with these approaches is that the 
number of time bins need to be chosen by the data analyst. On the one hand, a fine grouping 
of times may result in few (or none) events per interval, which then leads 
to instability in the Newton-Raphson procedure for estimation 
[@kalbfleisch2011statistical, Section 4.8]. 
On the other hand, a coarse grouping could potentially mask 
nonlinear trends in the hazard function.
